Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences

### 2019-02-15 draft.tex
improving MIP optimization

Let $\hat{v}$ be the compute profile computed according to the above procedure, and  $\hat{v}_j$ the $j$-th element (the linear order extending the partial order $p_j$ of the $j$-th voter).
The PMR can be written as:
\[ \PMR(x,y; \pprofile, W) = \sum_{j=1}^n w_{\hat{v}_i(y)} - w_{\hat{v}_i(x)} 
= \sum_{i=1}^m (\alpha_{i}^{y} - \alpha_{i}^{x}) w_i \]
where the coefficient $ (\alpha_{i}^{y} - \alpha_{i}^{x})$ is the number of times $y$ is in position $i$ minus the number of times $x$ is the in position $i$ in the rankings of the completed profile $\hat{v}$:
$\alpha_{i}^{x} = \sum_{j=1}^{n}  I[\hat{v}_{j}(x)=i]$.
%\[ \alpha_{i} = \sum_{j=1}^{n}  I[\hat{v}_{j}(y)=i] - I[\hat{v}_{j}(x)=i]\]
where $I$ is the indicator function.
Then, what we have to to compute pairwise max regret is to compute the solution of the following linear program :
\begin{align*}
\PMR(x,y; \pprofile,W) = \max_{\w} & \sum_{i=1}^m \alpha_i w_{i}\\
\text{ s.t. } & \w \in W \\
& \text{Equation } (\ref{eq:convexity})
\end{align*}
The result of the optimization is the value of the pairwise max regret $\PMR$.

@@

\medskip
Let $\hat{v}_i$ be the linear order extending $p_i$ according to the above procedure (described in \cite{Lu2011}).
Then PMR can be written as follows:
\begin{align*}
 &\PMR(x,y; \pprofile, W) =\\ 
 &\max_{\w \in W} \Big \{ \max_{v_j \in C(\succ_j^p)} [w_{v_j(y)} \!-\! w_{v_j(x)}]  \!+\!	 \sum_{j \in A^+ \cup  A^?} [w_{\hat{v}_j(y)} \!-\! w_{\hat{v}_j(x)}] \Big \} 
 \end{align*}

%We compute pairwise maximum regret by considering binary variables $\{ B_{i} \}_{i=1,\ldots,m}$ to represent optimization choices related to where to position the alternatives.
We introduce two set of binary variables $B^{+}_{i,j}$ and $B^{-}_{i,j}$  for each voter $i$ and each position $j$.
The variable $B_{i,j}^{+}$ encodes the fact that the alternative $y$ is placed in position $j$ in the ranking of agent $i$; while  $B_{i,j}^{-}$ encodes the position of the alternative $x$.

@@

The score of alternative $y$ can be written as $\sum_{i=1}^{n} \sum_{j=1}^{m} w_{j} B_{i,j}^{+}$.
Our objective function is now:
 \[ \max \sum_{i=1}^{n} \sum_{j=1}^{m} B_{i,j}^{+} w_j- B_{i,j}^{-}  w_j \]

We use integer programming enconding tricks in order to linearize the problem formulation.
We introduce yet another set of variables  $V_{i,j}^{+} $  and $V_{i,j}^{-}$ % the multiplicative terms by new variables.
and we introduce constraints so that $V_{i,j}^{+} = B^{+}_{i,j} w_j$.
To impose the desired behavior, we set $V_{l}^{j} \leq B^j $ and $V_{l}^{j} \leq w_l$.
%\begin{align*} 
%V_{l}^{j} \leq B^j \\
%V_{l}^{j} \leq w_l
%\end{align*}
%The expression $B^j_l w_l$ evaluates to $w_l$ if $I^j_l = 1$, and to $0$ otherwise.

@@

We therefore obtain the following mixed integer linear program:
\begin{align}
\max & \sum_{i=1}^n  \sum_{j=1}^m  V_{i,j}^{+} - V_{i,j}^{-}\\
\text{ s.t. } & w \in W  \\
& \sum_{i=t_{1}}^{t_{2}} B_{i,j}^{+} = 1 & \forall j \\
& V_{i,j}^{+} \leq B_{i,j}^{+}  & \forall i,j \\
& V_{i,j}^{+} \leq w_j & \forall i,j \\
& V_{i,j}^{-} \geq w + B_{i,j}^{-} - 1 & \forall i,j \\
& V_{i,j}^{-} \geq 0 & \forall i,j
\end{align} 
%We write $w \in W$ as a shourtcut to represent the constraints that the weights are chosen to be in the feasible set.

@@

In particular, we want to acquire constraints of the type:
\[ w_{j} - w_{j+1} \geq \lambda (w_{j+1} - w_{j+2}) \] 
for $j \in \{1,\ldots,m-2\}$.
This constraint relates the difference between the importance of positions $i$ and $i+1$ with the difference between positions $i+1$ and $i+2$.
While perhaps in some cases the decision maker may be able to state pferences of this kind, we argue that a more concrete way is better.
Notice that the constraint above is equivalent to $w_{i} - (\lambda+1) w_{i+1} + \alpha w_{i+2} \geq 0$.
%We conjecture that preferences of this kind are relatively easy to state
%(``do you think that the difference between the importance of position $i$ and $i+1$ is at least the difference between positions $i+1$ and $i+2$ ?'').
We present them a profile representing the situation described by the query and then deduce its answer from the selected winner. 
%We argue that it may be cognitively easier to ask to pick a winner from a profile, rather than answering an abstract question.
Therefore we generate profiles specifically in such a way that user answers gives us the kind of constraint that we want to acquire.
We provide a systematic way to construct such profiles.  %that reflects the situation outlined by the query for two alternatives and the others are not better than them.

@@

@@

\vspace{1cm}
{\bf Complete section}
\vspace{1cm}
***

@@

% We can stop elicitation when minimax regret is sufficiently small.
We test the elicitation protocols in randomly generated datasets and real datasets. % (sushi and iris).

We evaluate the proposed method using the following simulated protocol.
First of all, we randomly generate the true preferences of the users (i.e. the linear orders) and the weights associated with the committee's scoring rule.
For each of the elicitation strategy, we simulate an elicitation session.
%Queries are posed to simulated users, that answer accor
%We compare the performance  of the different strategies of Section \ref{sec:elicit} with respect to decrement in max regret and with respect to the real loss.
We measure the effectiveness of our strategies of Section \ref{sec:elicit} by examining regret reduction as a function of the number of queries.

>>>>

Let $\hat{v}$ be the compute profile computed according to the above procedure, and  $\hat{v}_j$ the $j$-th element (the linear order extending the partial order $p_j$ of the $j$-th voter).
The PMR can be written as:
\[ \PMR(x,y; \pprofile, W) = \sum_{j=1}^n w_{\hat{v}_i(y)} - w_{\hat{v}_i(x)} 
= \sum_{i=1}^m (\alpha_{i}^{y} - \alpha_{i}^{x}) w_i \]
where the coefficient $ (\alpha_{i}^{y} - \alpha_{i}^{x})$ is the number of times $y$ is in position $i$ minus the number of times $x$ is the in position $i$ in the rankings of the completed profile $\hat{v}$:
$\alpha_{i}^{x} = \sum_{j=1}^{n}  I[\hat{v}_{j}(x)=i]$ where $I$ is the indicator function.
%\[ \alpha_{i} = \sum_{j=1}^{n}  I[\hat{v}_{j}(y)=i] - I[\hat{v}_{j}(x)=i]\]
Then, what we have to to compute pairwise max regret is to compute the solution of the following linear program :
\begin{align*}
\PMR(x,y; \pprofile,W) = \max_{\w} & \sum_{i=1}^m (\alpha_{i}^{y} - \alpha_{i}^{x}) w_{i}\\
\text{ s.t. } & \w \in W \\
& \text{Equation } (\ref{eq:convexity})
\end{align*}
The result of the optimization is the value of the pairwise max regret $\PMR$.

@@

\medskip
Let $\hat{v}_i$ be the linear order extending $p_i$ according to the above procedure (described in \cite{Lu2011}).
Then PMR can be written as follows:
\begin{align*}
 &\PMR(x,y; \pprofile, W) =\\ 
 &\max_{\w \in W} \Big \{ \max_{v_j \in C(\succ_j^p)} [w_{v_j(y)} \!-\! w_{v_j(x)}]  \!+\!	 \sum_{j \in A^+ \cup  A^?} [w_{\hat{v}_j(y)} \!-\! w_{\hat{v}_j(x)}] \Big \} 
 \end{align*}
%Note that the second addendum inside the max do not depend on the choice of $\w$.
Consider the two addendum inside the outer ``max''. 
The first addendum is concern with positioning of alternatives $x$ and $y$ for voters in $A^{-}$for which we know that $x$ is preferred to $y$.
The second addendum is concerned with voters in $A^{+}$ and $A^{?}$.
We rewrite the second addendum as:
\begin{align*}
\sum_{j \in A^+ \cup  A^?} [w_{\hat{v}_j(y)} \!-\! w_{\hat{v}_j(x)}] 
= \sum_{i = 1}^{m} (\hat{\alpha}_{i}^{y} - \hat{\alpha}_{i}^{x}) w_{i}
\end{align*}
where $\hat{\alpha}_{i}^{x}$ is the number of times that $x$ is ranked  in position $i$  considering the profile $\hat{v}$ of voters in $A^+ \cup  A^?$.
%We compute pairwise maximum regret by considering binary variables $\{ B_{i} \}_{i=1,\ldots,m}$ to represent optimization choices related to where to position the alternatives.

We now show how to optimize $\PMR$.
In addition to variables $\{ w_{j} \}_{j=1,\ldots,m}$ (one for each position) we need to employ several additional decision variables.
We introduce two set of binary variables $B^{+}_{i,j}$ and $B^{-}_{i,j}$  for each position $i$ and for each voter $j$.
Variable $B_{i,j}^{+}$ encodes the fact that the alternative $y$ is placed in position $i$ in the ranking of agent $j$; while  $B_{i,j}^{-}$ encodes the same thing for alternative $x$.

@@

The score of alternative $y$ can be written as $\sum_{i = 1}^{m} \hat{\alpha}_{i}^{y} + \sum_{i=1}^{n} \sum_{j=1}^{m} w_{j} B_{i,j}^{+}$.
The objective function is now:
 \[ \max \sum_{i = 1}^{m} (\hat{\alpha}_{i}^{y} - \hat{\alpha}_{i}^{x}) w_{i}  \sum_{i=1}^{n} \sum_{j=1}^{m} B_{i,j}^{+} w_j- B_{i,j}^{-}  w_j \]

We use integer programming enconding tricks in order to linearize the problem formulation.
We introduce yet another set of variables  $V_{i,j}^{+} $  and $V_{i,j}^{-}$ % the multiplicative terms by new variables.
and we enforce that $V_{i,j}^{+} = B^{+}_{i,j} w_j$ by setting constraints $V_{l}^{j} \leq B^j $ and $V_{l}^{j} \leq w_l$.
%\begin{align*} 
%V_{l}^{j} \leq B^j \\
%V_{l}^{j} \leq w_l
%\end{align*}
%The expression $B^j_l w_l$ evaluates to $w_l$ if $I^j_l = 1$, and to $0$ otherwise.

@@

We therefore obtain the following mixed integer linear program:
\begin{align}
\max & \sum_{i=1}^m  [(\hat{\alpha}_{i}^{y} - \hat{\alpha}_{i}^{x}) w_{i}] +
 \sum_{i=1}^m \sum_{j=1}^n  [V_{i,j}^{+} - V_{i,j}^{-}]\\
\text{ s.t. } & \w \in W  \\
& \sum_{i=t_{1}}^{t_{2}} B_{i,j}^{+} = 1 & \forall j \\
& \sum_{i=t_{1}}^{t_{2}} B_{i,j}^{-} = 1 & \forall j \\
& V_{i,j}^{+} \leq B_{i,j}^{+}  & \forall i,j \\
& V_{i,j}^{+} \leq w_j & \forall i,j \\
& V_{i,j}^{-} \geq w + B_{i,j}^{-} - 1 & \forall i,j \\
& V_{i,j}^{-} \geq 0 & \forall i,j
\end{align} 
%We write $w \in W$ as a shourtcut to represent the constraints that the weights are chosen to be in the feasible set.

@@

In particular, we want to acquire constraints of the type:
\[ w_{j} - w_{j+1} \geq \lambda (w_{j+1} - w_{j+2}) \] 
for $j \in \{1,\ldots,m-2\}$, relating the difference between the importance of positions $i$ and $i+1$ with the difference between positions $i+1$ and $i+2$.
While perhaps in some cases the decision maker may be able to state preferences of this kind, we argue that a more concrete way is better.
Notice that the constraint above is equivalent to 
$w_{i} - (\lambda+1) w_{i+1} + \lambda w_{i+2} \geq 0$
and to $w_{i} + \lambda w_{i+2} \geq  (\lambda+1) w_{i+1}$.
%We conjecture that preferences of this kind are relatively easy to state
%(``do you think that the difference between the importance of position $i$ and $i+1$ is at least the difference between positions $i+1$ and $i+2$ ?'').
We present them a profile representing the situation described by the query and then deduce its answer from the selected winner. 
%We argue that it may be cognitively easier to ask to pick a winner from a profile, rather than answering an abstract question.
Therefore we generate profiles specifically in such a way that user answers gives us the kind of constraint that we want to acquire.
We provide a systematic way to construct such profiles.  %that reflects the situation outlined by the query for two alternatives and the others are not better than them.

@@

\vspace{1cm}
{\bf Complete section:} give a precise algorithm for completing the profile
\vspace{1cm}

@@

@@

% We can stop elicitation when minimax regret is sufficiently small.
We test the elicitation protocols in randomly generated datasets and real datasets. % (sushi and iris).
We evaluate the proposed method using the following simulated protocol.
First of all, we randomly generate the true preferences of the users (i.e. the linear orders) and the weights associated with the committee's scoring rule.
For each of the elicitation strategy, we simulate an elicitation session.
%Queries are posed to simulated users, that answer accor
%We compare the performance  of the different strategies of Section \ref{sec:elicit} with respect to decrement in max regret and with respect to the real loss.
We measure the effectiveness of our strategies of Section \ref{sec:elicit} by examining regret reduction as a function of the number of queries.
